### 2016-12-01 local-capability-machine.tex
Typo fixes and some small remarks...

Things to note:
\begin{itemize}
\item $\RegName$ contains $\pcreg$, but is otherwise a sufficiently
  large finite set.
\item Table~\ref{tab:permission-list} describes what all the permissions grant access to.
\item Figure~\ref{fig:perm-hier} shows the ordering of the permissions, i.e, the elements of $\Perms$.
\item Figure~\ref{fig:glob-hier} shows the ordering of $\local$ and $\glob$, i.e., the elements of $\Globals$.
\item The ordering of $\Perms \times \Globals$ is point wise.
\end{itemize}

@@

@@

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
\lau{ I think this is a conceptual question as we can make it work with either. We already trust malloc to give us a fresh piece of memory and not reuse it later on, so we already assume malloc to be somewhat trusted, so why not go all the way? }
\dominique{what does ``fetch the capability ...'' mean?}
\lau{ We don't know where the capability resides, but if it is in memory, then it will be loaded into a register. }
\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e. which do not need to be erased)?}
\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}

@@

\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
\label{lem:correctness-f1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\var{offset-link-flag},\mathtt{1f},\mathtt{f1}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\var{offset-link-flag}) \\
    m & \defeq \hs_{f1} \uplus 
               \hs_\flag \uplus                
               \ms_{\var{link}} \uplus 
               \hs_\adv \uplus 
               \ms_{\malloc} \uplus 
               \hs_{\var{frame}} 
  \end{align*}
where 
\begin{align*}
  &\dom(\hs_{f1}) = [\mathtt{f1}-\var{offset-link-flag},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\dom(\ms_\link) = [\link,\link+1]\\
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
\end{align*}
and
\begin{itemize}
\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f1}$ contains the code of $f1$.
\item $\ms_\flag = [flag \mapsto 0]$
\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
\item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
\end{itemize}
if 
  \[
    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
  \]
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

@@

\begin{lemma}
  \label{lem:correctness-f2}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\var{offset-link-flag},\mathtt{2f},\mathtt{f2}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f2}) = [\mathtt{f2}-\var{offset-link-flag},\mathtt{2f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item The first address of $\hs_{f2}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f2}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f2}$ contains the code of $f2$.
  \item $\ms_\flag = [flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
  \begin{multline*}
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
  \end{multline*}
Where 
\[
  m''  \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_\link \uplus 
        \hs_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk' \uplus
        \ms_{\var{frame}}
\]
Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
\begin{itemize}
\item The range of authority for 
  \begin{itemize}
  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
  \end{itemize}
  the two stack capabilities have the same permissions as $c_\stk$
\item $c_{f2}'$ points to the return address in f2's code.
\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
\end{itemize}

@@

which is indeed the case. Use this to apply the FTLR to get 

@@

    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$:\\ This region matches only this memory segment.

@@

@@

@@

      From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack). The restoration code remains unaltered, and so does the data on the stack, so the execution continues by popping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$. Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ instruction. Now the top word of the stack is popped to $r_1$, and thanks to the world satisfaction, we know that this word is 1. The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.

All in all, there are no changes to the memory, so we can conclude
\[
  m'' = \ms' \uplus \ms_f
\]
From Lemma~\ref{lem:priv-mono-like}, we can get $\ms''$ and $\ms_r$ such that
\begin{itemize}
\item $\ms' = \ms'' \uplus \ms_r$
\item $\heapSat[\ms'']{n}{\revokeTemp{W}}$
\end{itemize}
Now downwards closure of memory segment satisfaction gives us the desired result.
    \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
  \end{enumproof}
\end{enumproof}
From \ref{lem:f2-adv-er}, \ref{lem:f2-mem-sat}, and \ref{lem:f2-reg-rel} it follows that
\[
  \npair{(\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}\update{\pcreg}{c_\adv},\ms)} \in \observations(W)
\]
Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfuly gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
\begin{itemize}
\item $m' = \ms' \uplus\ms_r \uplus \ms_f$
\item $\heapSat[\ms']{n-i}{W'}$
\end{itemize}
The permanent region $\iota^\sta (\perma,\ms_\flag)$ ensures that $\ms_\flag$ is unchanged, so
\[
  m'(\flag) = 0
\]
\end{proof}

>>>>

Things to note:
\begin{itemize}
\item $\RegName$ contains $\pcreg$, but is otherwise a sufficiently
  large finite set.
\item Table~\ref{tab:permission-list} describes what all the permissions grant access to.
\item Figure~\ref{fig:perm-hier} shows the ordering of the permissions, i.e, the elements of $\Perms$.
\item Figure~\ref{fig:glob-hier} shows the ordering of $\local$ and $\glob$, i.e., the elements of $\Globals$.
\item The ordering of $\Perms \times \Globals$ is pointwise.
\end{itemize}

@@

\dominique{I see, but maybe writeAllowed could take $w$ and $c$ as an argument, rather than just $\perm$? It is now weird that some conditions for writing are in $\writeAllowed{}$, but not all.}

\dominique{factor out reoccuring conditions of the form $n_i = \rv_i$ or $n_i =
  \memreg(\rv_i)$ into a function that evaluates an $\rv$ using a system state?
  This would generalize nicely to the case where we would have additional
  addressing modes.}

\dominique{why doesn't store take an $\rv$?}

@@

\section{Macros}
In order to write readable example programs, we provide macros (macro-instructions) that can be implemented in terms of the instruction set given in the formalisation.
\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
\lau{ I think this is a conceptual question as we can make it work with either. We already trust malloc to give us a fresh piece of memory and not reuse it later on, so we already assume malloc to be somewhat trusted, so why not go all the way? }
\lau{Agreed.}

\dominique{what does ``fetch the capability ...'' mean?}
\lau{ We don't know where the capability resides, but if it is in memory, then it will be loaded into a register. }
\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e. which do not need to be erased)?}
\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}

@@

\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
\label{lem:correctness-f1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\var{offset-link-flag},\mathtt{1f},\mathtt{f1}) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\var{offset-link-flag}) \\
    m & \defeq \hs_{f1} \uplus 
               \hs_\flag \uplus                
               \ms_{\var{link}} \uplus 
               \hs_\adv \uplus 
               \ms_{\malloc} \uplus 
               \hs_{\var{frame}} 
  \end{align*}
where 
\begin{align*}
  &\dom(\hs_{f1}) = [\mathtt{f1}-\var{offset-link-flag},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\dom(\ms_\link) = [\link,\link+1]\\
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
\end{align*}
and
\begin{itemize}
\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f1}$ contains the code of $f1$. \dominique{why not write this in symbols?}
\item $\ms_\flag = [flag \mapsto 0]$
\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
\item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
\end{itemize}
if 
  \[
    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
  \]
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

@@

\begin{lemma}
  \label{lem:correctness-f2}
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-(\var{offset-link-flag}),\mathtt{2f},\mathtt{f2}) \\
    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
    \reg & \in \Regs \\
    m & \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_{\var{link}} \uplus 
        \hs_\adv \uplus 
        \ms_{\malloc} \uplus 
        \ms_{\var{stk}} \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{align*}
    &\dom(\hs_{f2}) = [\mathtt{f2}-\var{offset-link-flag},\mathtt{2f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link+1]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
  \end{align*}
  and
  \begin{itemize}
  \item The first address of $\hs_{f2}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f2}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f2}$ contains the code of $f2$.
  \item $\ms_\flag = [flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
  \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
  \end{itemize}
  if 
  \[
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
  \begin{multline*}
    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
  \end{multline*}
Where 
\[
  m''  \defeq \hs_{f2} \uplus 
        \hs_\flag \uplus                
        \ms_\link \uplus 
        \hs_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk' \uplus
        \ms_{\var{frame}}
\]
Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
\begin{itemize}
\item The range of authority for 
  \begin{itemize}
  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
  \end{itemize}
  the two stack capabilities have the same permissions as $c_\stk$
\item $c_{f2}'$ points to the return address in f2's code.
\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
\end{itemize}

@@

which is indeed the case\dominique{by some lemma?}. Use this to apply the FTLR to get 

@@

    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in
      H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$ \dominique{apply to world?}:\\ This region matches only this memory segment.

@@

\dominique{Factor out a reusable lemma?}

@@

      \dominique{part of a phrase missing here?}

@@

      From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack). The restoration code remains unaltered, and so does the data on the stack, so the execution continues by popping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$. Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ macro-instruction. Now the top word of the stack is popped to $r_1$, and thanks to the world satisfaction, we know that this word is 1. The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.
      \dominique{Maybe you can factor out some sort of reusable antireduction lemma
        for this sort of argument?}

All in all, there are no changes to the memory, so we can conclude
\[
  m'' = \ms' \uplus \ms_f
\]
From Lemma~\ref{lem:priv-mono-like}, we can get $\ms''$ and $\ms_r$ such that
\begin{itemize}
\item $\ms' = \ms'' \uplus \ms_r$
\item $\heapSat[\ms'']{n}{\revokeTemp{W}}$
\end{itemize}
Now downwards closure of memory segment satisfaction gives us the desired result.
    \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
  \end{enumproof}
\end{enumproof}
From \ref{lem:f2-adv-er}, \ref{lem:f2-mem-sat}, and \ref{lem:f2-reg-rel} it follows that
\[
  \npair{(\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}\update{\pcreg}{c_\adv},\ms)} \in \observations(W)
\]
Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfully gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
\begin{itemize}
\item $m' = \ms' \uplus\ms_r \uplus \ms_f$
\item $\heapSat[\ms']{n-i}{W'}$
\end{itemize}
The permanent region $\iota^\sta (\perma,\ms_\flag)$ ensures that $\ms_\flag$ is unchanged, so
\[
  m'(\flag) = 0
\]
\end{proof}
